Linear vertical shear instability in astrophysical disks: a semi-numerical study

### 2014-12-18 appendix.tex
added appendix b, edited first numerical stuff

>>>>

\section{Improving the nearly-Keplerian approximation for adiabatic
  disks}\label{adia_improve}
In \S\ref{analytic_adia} we made the replacement $D\to\Omega_k^2$
before combining Eq. \ref{adia_iso1}---\ref{adia_iso2} to derive
Eq. \ref{adia_iso3}. This procedure ignores the vertical dependence of
$D$. We show here that this has no significant consequence for thin
disks.

Differentiating Eq. \ref{adia_iso1} properly gives
\begin{align}
  \left(D + \gamma k_x^2 c_s^2\right)\frac{d\Delta}{dz}=& D
  \frac{d^2\delta v_z}{dz^2} + k_x^2c_s^2\left[\frac{\gamma}{\left(D + \gamma k_x^2 c_s^2\right)}\frac{dD}{dz} + \ii
    \frac{d\ln\rho}{dz}\left(\ii +
      \frac{q}{k_xr}\right)\right]\frac{d\delta v_z}{dz}\notag\\
  &+ \ii k_x^2 c_s^2 \left(\ii +
      \frac{q}{k_xr}\right)\left[\frac{d^2\ln\rho}{dz^2} -
      \frac{1}{\left(D + \gamma
          k_x^2c_s^2\right)}\frac{dD}{dz}\frac{d\ln\rho}{dz}\right]\delta
    v_z. \label{adia_diso1}
\end{align}
The terms that were ignored in deriving Eq. \ref{adia_iso3} are those
proportional to $dD/dz$ in Eq. \ref{adia_diso1}. We estimate the
importance of these terms by noting that 
\begin{align}
  \frac{dD}{dz}\equiv \frac{d\kappa^2}{dz} \simeq \frac{d\Omega^2}{dz}
  = - \frac{d\ln\rho}{dz}\frac{qc_s^2}{r^2},
\end{align}
for thin disks. Then Eq. \ref{adia_diso1} becomes
\begin{align}
\left(D + \gamma k_x^2 c_s^2\right)\frac{d\Delta}{dz}=& D
  \frac{d^2\delta v_z}{dz^2} - k_x^2c_s^2\frac{d\ln\rho}{dz}\left[ 1 - 
      \frac{\ii q}{k_xr}  +  \frac{\gamma q c_s^2}{r^2\left(D +
        \gamma k_x^2 c_s^2\right)}\right]\frac{d\delta
    v_z}{dz}\notag\\ 
  &- k_x^2 c_s^2 \left(1 - 
      \frac{\ii q}{k_xr}\right)\left[\frac{d^2\ln\rho}{dz^2} + 
      \frac{qc_s^2}{r^2\left(D + \gamma
          k_x^2c_s^2\right)}\left(\frac{d\ln\rho}{dz}\right)^2\right]\delta
    v_z. \label{adia_diso2}
\end{align}
Since $D\sim \Omega_k^2$ in the low-frequency limit, we 
see from Eq. \ref{adia_diso2} that the neglected terms are
$O(\epsilon^2)$. We can make this explicit by combining
Eq. \ref{adia_diso1}---\ref{adia_diso2} and Eq. \ref{adia_iso2}, then
set $D\to\Omega_k^2$. In terms of non-dimensional variables, the result
is  
\begin{align}
   &\delta v_z ^{\prime\prime} + \left(1 + \ii\epsilon q\hat{k} -
    \frac{\gamma q \epsilon^2\hat{k}^2}{1+\gamma
      \hat{k}^2}\right)\ln\rho^{\prime}\delta v_z^\prime +
  \left[\left(\frac{1}{\gamma} + \ii \epsilon q
      \hat{k}\right)\ln\rho^{\prime\prime} - \hat{k}^2\left(1 -
      \frac{\ii\epsilon
        q}{\hat{k}}\right)\left(\frac{\gamma-1}{\gamma} +
      \frac{q\epsilon^2}{1+\gamma\hat{k}^2}\right)\ln\rho^{\prime
      2}\right]\delta v_z \notag\\&=
  -\hat{\sigma}^2\left(\frac{1}{\gamma} + \hat{k}^2\right)\delta v_z\label{adia_diso3}.
\end{align}   
Eq. \ref{adia_diso3} differs from Eq. \ref{adia_iso3} by terms
proportional to $\epsilon^2$. For a thin disk, $\epsilon\ll1$, so
neglecting these terms has no qualitative effect on the discussion in
\S\ref{analytic_adia}.

### 2014-12-18 linear_adia.tex
added appendix b, edited first numerical stuff

\subsection{Adiabatic perturbations} 
We now extend the above discussion to general $\gamma$, but in the
adiabatic limit ($t_c\to\infty$). We eliminate variables from 
Eq. \ref{lin_mass}---\ref{lin_energy} in favor of $\delta v_z$ 
and $\Delta \equiv \nabla\cdot\delta\bm{v}$ to obtain
\begin{align}
  &\Delta\left(1 + \frac{\gamma c_s^2 k_x^2}{D}\right) = \frac{d\delta
    v_z}{dz} + \frac{\ii k_x^2}{D}\left(\ii c_s^2 \frac{d\ln\rho}{dz}
    - \frac{r}{k_x}\frac{d\Omega^2}{dz}\right)\delta v_z,\label{adia_iso1}\\
  & -\frac{\sigma^2}{c_s^2}\delta v_z = \gamma \frac{d\Delta}{dz} +
  \frac{d}{dz}\left(\frac{d\ln\rho}{dz}\delta v_z\right) +
  \left(\gamma-1\right)\frac{d\ln\rho}{dz} \Delta.\label{adia_iso2}
\end{align}
%{\bf note: incompressible mode, set delta to zero. only works for
% constant gravity.}

We make the low-frequency approximation and the replacement
$D\to \Omega_k^2$ in Eq. \ref{adia_iso1}---\ref{adia_iso2}. For the
analysis to be carried out, the latter is a more limiting
approximation than for isothermal perturbations because we make the
replacement before combining the pair of equations above.

Here, we also use the fact that in the
global disk, 
\begin{align}
  r\frac{\p \Omega^2}{\p z} = -\frac{\p \ln\rho}{\p z} \frac{qc_s^2}{r},
\end{align}
for the power-law prescription of the midplane sound-speed for
$\Gamma=1$.

Making the above substitutions and combining Eq. \ref{adia_iso1}---
Eq. \ref{adia_iso2} we obtain, in terms of dimensionless variables defined
previously, an equation for $\delta v_z$, 
\begin{align}
  0 =& \dd v_z^{\prime\prime} + \left(1 + \ii \epsilon q
    \hat{k}\right)\left(\ln\rho^{\prime}\delta v_z\right)^\prime\notag\\
  &+\left\{\hat{\sigma}^2\left(\frac{1}{\gamma}+\hat{k}^2\right)\right.\notag\\
  &\phantom{0=}\left.-\left(\frac{\gamma-1}{\gamma}\right)\left[\ln\rho^{\prime\prime}+\hat{k}^2\left(1-\frac{\ii\epsilon  
          q}{\hat{k}}\right)\ln\rho^{\prime 2}\right]\right\}\delta v_z.\label{adia_iso3}
\end{align}

@@

Note that the term with $\ln\rho^{\prime 2}$ in the integrand (third
term on the RHS) represents stabilization by the vertical entropy
gradient when there is no vertical shear. It gives an imaginary
contribution to $\hat{\sigma}$ when $q\neq0$, but this is a factor
$|\epsilon/\hat{k}|$ smaller than the real (stabilizing) contribution,
since $\epsilon\ll 1$ for a thin disk and we (usually) consider
$\hat{k}\gg1$. Thus, instability is expected to associated with the 
last term on the RHS of Eq. \ref{adia_integral}.

\subsubsection{Explicit solutions in the thin-disk limit}
We can construct  solutions in the thin-disk limit, in which 
$\ln(\rho/\rho_0) \simeq -\zhat^2/2$. Eq. \ref{adia_iso3} becomes   
\begin{align}
  0 = \delta v_z^{\prime\prime} - \zhat A \delta v_z^\prime + \left(B
    - C \zhat^2\right)\delta v_z,
\end{align}
with
\begin{align}
  &A \equiv 1 + \ii \epsilon q \hat{k},\\
  &B \equiv \hat{\sigma}^2\left(\frac{1}{\gamma} + \hat{k}^2\right) -
  \left(\frac{1}{\gamma} + \ii \hat{k} \epsilon q\right)\\
  &C \equiv \frac{\left(\gamma-1\right)}{\gamma}\hat{k}^2\left(1 - \frac{\ii
      \epsilon q}{\hat{k}}\right).\label{adia_thin}
\end{align}
Let
\begin{align}
  \delta v_z(\zhat) =
  g(\zhat)\exp\left(\frac{\alpha\zhat^2}{2}\right), \label{adia_ansatz}
\end{align}
where $\alpha$ is a constant to be chosen for convenience. Inserting
Eq. \ref{adia_ansatz} into Eq. \ref{adia_thin} gives
\begin{align}
  0 = g^{\prime\prime} - \hat{z}\left(A - 2\alpha\right)g^\prime + \left(B +
    \alpha\right)g
  +\left(\alpha^2 - \alpha A - C\right)\zhat^2 g.
\end{align}
We choose $\alpha$ to make the coefficient of $\zhat^2g$
vanish. Furthermore, we require 
the vertical kinetic energy density $\rho|\delta v_z|^2$ to remain
finite. Then assuming $g(\zhat)$ is a polynomial, we require 
\begin{align}
  \real\alpha < \frac{1}{2}. 
\end{align}
We therefore choose 
\begin{align}
  \alpha = \frac{1}{2}\left(A - \sqrt{A^2 + 4C}\right). 
\end{align}

We now have an equation in the same form as Eq. \ref{iso_ode3}:
\begin{align}
  0 = g^{\prime\prime} - \hat{z}\left(A - 2\alpha\right)g^\prime +
  \left(B + \alpha\right)g.
\end{align}
We seek polynomial solutions 
\begin{align}
  g(\zhat) = \sum_{m=0}^M b_m \zhat^m,
\end{align}
% We again seek power-series solutions 
% \begin{align}
%   g(\zhat) = \sum_{m=0}^\infty b_m \zhat^m,
% \end{align}
% which requires
% \begin{align}
%   (m+2)(m+1)b_{m+2} + \left[\left(2\alpha - A\right)m + B + \alpha\right]=0.
% \end{align}
which requires
\begin{align}
  B + \alpha = M\left(A-2\alpha\right).
\end{align}
Inserting the definition of $B$ we get
\begin{align}
  &\hat{\sigma}^2\left(\frac{1}{\gamma}+\hat{k}^2\right) \notag\\ &=
  \left(\frac{1}{\gamma}-\frac{1}{2}\right) +\frac{\ii}{2}\epsilon q
  \hat{k} + \frac{1}{2}\left(2M+1\right)\left(A^2 + 4C\right)^{1/2}\label{adia_disp}. 
\end{align}
%Recalling that $B$ depends $\hat{\sigma}$, Eq. \ref{adia_disp} gives
%the eigenfrequency as a function of disk paramters.

\subsubsection{The case of $\gamma=2$}
In principle, we can solve Eq. \ref{adia_disp} for the mode
frequencies and growth rates. This is complicated in 
general, but simplifies for $\gamma=2$, for which the quantity
$A^2+4C$ is real. In this case we find
\begin{align}
  \hat{\nu}^2 =
  \frac{1}{2}\frac{(2M+1)}{\sqrt{1+2\hat{k}^2}}&\left[
    \sqrt{1-\frac{4M(M+1)\hat{k}^2\epsilon^2q^2}{(2M+1)^2(1+2\hat{k}^2)}}\right.\notag\\
  &\phantom{=}
  \left.- \sqrt{1 - \frac{\hat{k}^2\epsilon^2q^2}{(1+2\hat{k}^2)}}\right].
\end{align}
In the limit of a thin disk $\epsilon\to0$ or weak shear $|q|\to0$, we
have 
\begin{align}
  \hat{\nu}^2 = \frac{\epsilon^2
    q^2}{4(2M+1)}\frac{\hat{k}^2}{(1+2\hat{k}^2)^{3/2}},\label{gam2_growth}
\end{align}
and the maximum growth rate occurs at $\hat{k}=\hat{k}_\mathrm{opt}=1$,  
\begin{align}
  \mathrm{max}(\hat{\nu}) &= \frac{|\epsilon
    q|}{2(3^{3/4}\sqrt{2M+1})}\label{gam2_max_growth}.
  % & \leq \frac{|\epsilon
  % q|}{2\times3^{3/4}}\notag.
\end{align}
Unlike for isothermal perturbations, Eq. \ref{gam2_max_growth} show
that growth rates decrease with $M$, and there exists a maximum
possible growth rate at $M=0$. This is despite the fact that in the
thin-disk approximation, for which $d\Omega^2/dz\propto z$, there is
infinite vertical shear available.

The acceptable solution for $\alpha$ is
\begin{align}
  \alpha = \frac{1}{2}\left[\left(1+\ii\epsilon q \hat{k}\right) -
    \sqrt{1 + \hat{k}^2\left(2-\epsilon^2 q^2\right)}\right]. 
\end{align}
Since $\epsilon\ll 1$, $\real\alpha<0$, so the vertical velocity
eventually decays for large heights. This is unlike for isothermal
perturbations, in which perturbation magnitudes increase with $|z|$.

@@

In Fig. \ref{adia_growth} we plot the growth rate $|\hat{\nu}|$ as a
function of the radial wavenumber $\hat{k}$, for a range of adiabatic
indices, using Eq. \ref{adia_disp}. The presence of a positive
vertical entropy gradient is strongly stabilizing. This is because the
vertical shear grows as $z$ away from the midplane (and has a maximum
when the thin-disk approximation is relaxed), but the
square of the bouyancy frequency, which is stabilizing, grows as $z^2$
away from the midplane. The entropy $P/\rho^\gamma$ formally diverges
for $z\to\infty$ even without the thin-disk approximation.

>>>>

\subsection{Adiabatic perturbations}\label{analytic_adia}
We now extend the above discussion to general $\gamma$, but in the
adiabatic limit ($t_c\to\infty$). We eliminate variables from 
Eq. \ref{lin_mass}---\ref{lin_energy} in favor of $\delta v_z$ 
and $\Delta \equiv \nabla\cdot\delta\bm{v}$ to obtain
\begin{align}
  &\Delta\left(1 + \frac{\gamma c_s^2 k_x^2}{D}\right) = \frac{d\delta
    v_z}{dz} + \frac{\ii k_x^2}{D}\left(\ii c_s^2 \frac{d\ln\rho}{dz}
    - \frac{r}{k_x}\frac{d\Omega^2}{dz}\right)\delta v_z,\label{adia_iso1}\\
  & -\frac{\sigma^2}{c_s^2}\delta v_z = \gamma \frac{d\Delta}{dz} +
  \frac{d}{dz}\left(\frac{d\ln\rho}{dz}\delta v_z\right) +
  \left(\gamma-1\right)\frac{d\ln\rho}{dz} \Delta.\label{adia_iso2}
\end{align}
%{\bf note: incompressible mode, set delta to zero. only works for
% constant gravity.}

We make the low-frequency approximation and the replacement 
$D\to \Omega_k^2$ in Eq. \ref{adia_iso1}---\ref{adia_iso2}. For
simplicity we do this before combining these equations later. We  
show in Appendix \ref{adia_improve} that making these approximations
after combining  Eq. \ref{adia_iso1}---\ref{adia_iso2} adds unecessary
complexity for thin disks. Here, we also use the fact that in the
global disk, 
\begin{align}
  r\frac{\p \Omega^2}{\p z} = -\frac{\p \ln\rho}{\p z} \frac{qc_s^2}{r},
\end{align}
for the power-law prescription of the midplane sound-speed for
$\Gamma=1$.

Making the above substitutions and eliminating $\Delta$ between Eq. \ref{adia_iso1}---
Eq. \ref{adia_iso2} we obtain, in terms of dimensionless variables defined
previously, an equation for $\delta v_z$, 
\begin{align}
  0 =& \dd v_z^{\prime\prime} + \left(1 + \ii \epsilon q
    \hat{k}\right)\left(\ln\rho^{\prime}\delta v_z\right)^\prime\notag\\
  &+\left\{\hat{\sigma}^2\left(\frac{1}{\gamma}+\hat{k}^2\right)\right.\notag\\
  &\phantom{0=}\left.-\left(\frac{\gamma-1}{\gamma}\right)\left[\ln\rho^{\prime\prime}+\hat{k}^2\left(1-\frac{\ii\epsilon  
          q}{\hat{k}}\right)\ln\rho^{\prime 2}\right]\right\}\delta v_z.\label{adia_iso3}
\end{align}

@@

Note that when $\gamma>1$ and $q=0$, the term with $\ln\rho^{\prime
  2}$ in the integrand (third term on the RHS) represents
stabilization by the vertical entropy gradient (real bouyancy frequency). 
It gives an imaginary contribution to $\hat{\sigma}$ when $q\neq0$,
but this is a factor $|\epsilon/\hat{k}|$ smaller than the real
(stabilizing) contribution, 
since $\epsilon\ll 1$ for a thin disk and we (usually) consider
$\hat{k}\gg1$. Thus, instability is expected to be associated with the 
last term on the RHS of Eq. \ref{adia_integral}.

\subsubsection{Explicit solutions in the thin-disk limit}
We can construct solutions in the thin-disk limit, in which 
$\ln(\rho/\rho_0) \simeq -\zhat^2/2$. Eq. \ref{adia_iso3} becomes   
\begin{align}
  0 = \delta v_z^{\prime\prime} - \zhat A \delta v_z^\prime + \left(B
    - C \zhat^2\right)\delta v_z,
\end{align}
with
\begin{align}
  &A \equiv 1 + \ii \epsilon q \hat{k},\\
  &B \equiv \hat{\sigma}^2\left(\frac{1}{\gamma} + \hat{k}^2\right) -
  \left(\frac{1}{\gamma} + \ii \hat{k} \epsilon q\right)\\
  &C \equiv \frac{\left(\gamma-1\right)}{\gamma}\hat{k}^2\left(1 - \frac{\ii
      \epsilon q}{\hat{k}}\right).\label{adia_thin}
\end{align}
Let
\begin{align}
  \delta v_z(\zhat) =
  g(\zhat)\exp\left(\frac{\alpha\zhat^2}{2}\right), \label{adia_ansatz}
\end{align}
where $\alpha$ is a constant to be chosen for convenience. Inserting
Eq. \ref{adia_ansatz} into Eq. \ref{adia_thin} gives
\begin{align}
  0 = g^{\prime\prime} - \hat{z}\left(A - 2\alpha\right)g^\prime + \left(B +
    \alpha\right)g
  +\left(\alpha^2 - \alpha A - C\right)\zhat^2 g.
\end{align}
We choose $\alpha$ to make the coefficient of $\zhat^2g$
vanish, and impose the vertical kinetic energy density
$\rho|\delta v_z|^2$ to remain finite. Then assuming $g(\zhat)$ is a
polynomial, we require  
\begin{align}
  \real\alpha < \frac{1}{2}. 
\end{align}
We therefore choose 
\begin{align}
  \alpha = \frac{1}{2}\left(A - \sqrt{A^2 + 4C}\right). 
\end{align}

We now have an equation in the same form as Eq. \ref{iso_ode3}:
\begin{align}
  0 = g^{\prime\prime} - \hat{z}\left(A - 2\alpha\right)g^\prime +
  \left(B + \alpha\right)g.
\end{align}
We seek polynomial solutions 
\begin{align}
  g(\zhat) = \sum_{m=0}^M b_m \zhat^m,
\end{align}
% We again seek power-series solutions 
% \begin{align}
%   g(\zhat) = \sum_{m=0}^\infty b_m \zhat^m,
% \end{align}
% which requires
% \begin{align}
%   (m+2)(m+1)b_{m+2} + \left[\left(2\alpha - A\right)m + B + \alpha\right]=0.
% \end{align}
which requires
\begin{align}
  B + \alpha = M\left(A-2\alpha\right).
\end{align}
Inserting the definition of $B$ we get
\begin{align}
  &\hat{\sigma}^2\left(\frac{1}{\gamma}+\hat{k}^2\right) \notag\\ &=
  \left(\frac{1}{\gamma}-\frac{1}{2}\right) +\frac{\ii}{2}\epsilon q
  \hat{k} + \frac{1}{2}\left(2M+1\right)\left(A^2 + 4C\right)^{1/2}\label{adia_disp}. 
\end{align}
This gives the eigenfrequency $\hat{sigma}$ and completes the solution
description.

%Recalling that $B$ depends $\hat{\sigma}$, Eq. \ref{adia_disp} gives
%the eigenfrequency as a function of disk paramters.

\subsubsection{The case of $\gamma=2$}
In principle, we can solve Eq. \ref{adia_disp} for the mode
frequencies and growth rates as a function of the disk parameters and
the radial wavenumber. This is complicated in 
general, but simplifies for $\gamma=2$, for which the quantity
$A^2+4C$ is real. In this case we find
\begin{align}
  \hat{\nu}^2 =
  \frac{1}{2}\frac{(2M+1)}{\sqrt{1+2\hat{k}^2}}&\left[
    \sqrt{1-\frac{4M(M+1)\hat{k}^2\epsilon^2q^2}{(2M+1)^2(1+2\hat{k}^2)}}\right.\notag\\
  &\phantom{=}
  \left.- \sqrt{1 - \frac{\hat{k}^2\epsilon^2q^2}{(1+2\hat{k}^2)}}\right].
\end{align}
In the limit of a thin disk $\epsilon\to0$ or weak shear $|q|\to0$, we
have 
\begin{align}
  \hat{\nu}^2 = \frac{\epsilon^2
    q^2}{4(2M+1)}\frac{\hat{k}^2}{(1+2\hat{k}^2)^{3/2}},\label{gam2_growth}
\end{align}
and the maximum growth rate occurs at $\hat{k}=\hat{k}_\mathrm{opt}=1$,  
\begin{align}
  \mathrm{max}(\hat{\nu}) &= \frac{|\epsilon
    q|}{2(3^{3/4}\sqrt{2M+1})}\label{gam2_max_growth}.
  % & \leq \frac{|\epsilon
  % q|}{2\times3^{3/4}}\notag.
\end{align}
Unlike isothermal perturbations, Eq. \ref{gam2_max_growth} shows that
there is a maximum possible growth rate for $M=0$, and growth rates decrease with increasing $M$. This
is despite the fact that in the thin-disk approximation 
$d\Omega^2/dz\propto z$, so there is infinite vertical shear
available.

The acceptable solution for $\alpha$ when $\gamma=2$ is
\begin{align}
  \alpha = \frac{1}{2}\left[\left(1+\ii\epsilon q \hat{k}\right) -
    \sqrt{1 + \hat{k}^2\left(2-\epsilon^2 q^2\right)}\right]. 
\end{align}
Since $\epsilon\ll 1$, $\real\alpha<0$, so the vertical velocity
eventually decays for large heights. For $\hat{k}\gg 1$ the
perturbation decays rapidly away from the midplane. This is again
unlike isothermal perturbations, whose magnitude increase with $|z|$.

%gauss decay means boundary conditions have no influence, all perts go
%to zero

@@

In Fig. \ref{adia_growth} we plot the eigenfrequencies as a
function of the radial wavenumber $\hat{k}$, for a range of adiabatic
indices, using Eq. \ref{adia_disp}. The low-frequency approximation
is questionable for $\hat{k}\lesssim4$ since
$|\hat{\sigma}|^2\gtrsim0.2$. For large $\hat{k}$ the presence of a
positive vertical entropy gradient is strongly stabilizing. This is
because the vertical shear grows as $z$ away from the midplane (and
has a maximum when the thin-disk approximation is relaxed), but square
of the bouyancy frequency, which is stabilizing, grows as $z^2$ away
from the midplane. %The entropy $P/\rho^\gamma$ formally diverges   
%for $z\to\infty$ even without the thin-disk approximation.

### 2014-12-18 numerical.tex
minor updates
added appendix b, edited first numerical stuff

\section{Numerical solutions}
In this section we numerically solve the linear stability problem
corresponding to Eq. \ref{iso_ode1}. We thus relax the thin-disk
assumption, using the full expressions for $\rho$, $\Omega$ and
$\kappa$, so that we may consider vertical domains up to $z\sim r$.
However, we retain the low-frequency approximation
$|\sigma^2|\ll\kappa^2$ to obtain a linear eigenvalue problem. This
level of approximation lies in between the full problem 
Eq. \ref{iso_ode}, solved by \cite{mcnally14}, and the thin-disk
limit Eq. \ref{iso_ode3}, solved by \cite{nelson13}.

We solve Eq. \ref{iso_ode1} using a pseudo-spectral method by
expanding $W$ in Chebyshev polynomials $T_l$ up to $l=512$ and
discretizing the equation on a grid with
$\zmax=10H_\mathrm{iso}$. This procedure converts the linear problem
to a standard matrix eigenvalue problem. %In the low-frequency
%approximation the eigenvalue is $\sigma^2$.

\subsection{Fiducial calculation}
We first validate the low-frequency approximation by solving a 
fiducial problem with $q=-1$, $p=-1.5$, $\epsilon=0.1$ and $k_x =
200\pi/r$ with solid boundaries. This is the setup considered by
\cite{mcnally14} who solved the full problem.

Fig. \ref{lowfreq_eigen} shows the eigenvalues $\sigma = \omega +
\ii\nu$. This plot is effectively identical to that in
\cite{mcnally14}. The low-frequency approximation reproduces the
three solution branches defined by \citep{nelson13}: corrugation modes
with odd symmetry, breathing modes with even symmetry, and surface
modes which correspond to the nearly-vertical locus in
Fig. \ref{lowfreq_eigen}.

@@

We plot in Fig. \ref{lowfreq_eigenfunc} the structurally simplest
eigenfunction --- the fundamental corrugation mode corresponding to
the eigenvalue with $\mathrm{min}|\sigma|$. The perturbation $W$ is
roughly proportional to $z$, except close the boundaries where it
flattens because $\delta v_z=dW/dz=0$ is imposed there.

For $k_x = 200\pi/r$ and $\epsilon = 0.1$, the dimensionless
wavenumber $\hat{k} = k_xH_\mathrm{iso}=20\pi$. Then the expected
growth rate according to Eq. \ref{simple_growth} with $M=1$ is
\begin{align*}
  \nu = 0.2606\epsilon\Omega_k,
\end{align*}
as obtained numerically (Fig. \ref{lowfreq_eigen}). This agreement is
surpsingly good, given that Eq. \ref{simple_growth} assumes the
thin-disk approximation and imposes a different boundary condition to
that in the numerical calculations.

\begin{figure}
  \includegraphics[width=\linewidth]{figures/eigenvector_iso}
  \caption{Eigenfunction of the fundamental corrugation mode,
    corresponding to the 
    bottom-left eigenvalue displayed in Fig. \ref{lowfreq_eigen}. The
    solid and dashed lines are $\real W$ and $\imag W$, respectively. 
    \label{lowfreq_eigenfunc}
  }
\end{figure}

>>>>

\section{Numerical solutions in the low frequency limit}
In this section we numerically solve the linear stability problem.  
We relax the thin-disk assumption, using the full, non-Keplerian
expressions of $\rho$, $\Omega$ and $\kappa$, so that we may consider
vertical domains up to $z\sim r$. However, we retain the low-frequency
approximation $|\sigma^2|\ll\kappa^2$ to obtain a linear eigenvalue
problem, which limits us to consider large radial wavenumbers
$k_x$. We will relax this in the next section when considering
moderate values of $k_x$. % This level of approximation lies in between
% the full problem Eq. \ref{iso_ode}, solved by \cite{mcnally14}, and
% the thin-disk limit Eq. \ref{iso_ode3}, solved by \cite{nelson13}.

%We will solve different versions of the linearized equations depending
%on the disk properties, but in all cases we use

\subsection{Vertically isothermal disks with isothermal perturbations}   
We first validate the low-frequency approximation by re-producing
results from \cite{mcnally14} who solved the full problem for
vertically isothermal disks ($\Gamma=1$) subject to isothermal
perturbations ($\gamma=1$). The parameters are $q=-1$,
$p=-1.5$, $\epsilon=0.1$ and $k_x = 200\pi/r$ (or $\hat{k} = 20\pi$),
with solid vertical boundaries.

For this problem we solve Eq. \ref{iso_ode1}, using a pseudo-spectral
method by expanding $W$ in Chebyshev polynomials $T_l$ up to $l=512$
and discretizing the equation on a grid with
$\zmax=10H_\mathrm{iso}$. This procedure converts the linear problem
to a standard matrix eigenvalue problem.

Fig. \ref{lowfreq_eigen} shows the eigenvalues $\sigma = \omega +
\ii\nu$. This plot is effectively identical to that in
\cite{mcnally14}. The fundamental VSI mode, with one node in $W$, has
the smallest $|\sigma|$ and is shown in Fig. \ref{lowfreq_eigenfunc}.  
The expected growth rate from Eq. \ref{simple_growth} with 
$L=1$ is  
\begin{align*}
  \nu = 0.2606\epsilon\Omega_k,
\end{align*}
as obtained numerically. This agreement is surpsingly good, given that
Eq. \ref{simple_growth} assumes the thin-disk approximation and
imposes a different boundary condition to that in the numerical
calculation. The eigenfunction $W$ is qualitatively proportional to
$z$, as expected for the fundamental mode according to explicit
solutions discussed in \S\ref{iso_poly}.

@@

\begin{figure}
  \includegraphics[width=\linewidth]{figures/eigenvector_iso}
  \caption{Eigenfunction of the fundamental VSI,
    corresponding to the bottom-left eigenvalue displayed in
    Fig. \ref{lowfreq_eigen} (smallest $|\sigma|$). The
    solid and dashed lines are $\real W$ and $\imag W$, respectively. 
    \label{lowfreq_eigenfunc}
  }
\end{figure}

\subsection{Stabilization of the VSI by a positive vertical entropy
  gradient}
